Drivers of tree regeneration remain challenging to be captured by models of

### 2022-12-02 manuscript/manuscriptRW.Rmd
a failed attempt: trying to write conclusion in the manuscript

---
title: Drivers of tree regeneration remain challenging to be captured by models of
  forest dynamics
author:
- Olalla Díaz-Yáñez,
- Yannek Käber,
- Mats Mahnken,
- Samuel Fischer,
- Xavier Morin,
- Manfred Lexer,
- Christian Hochauer,
- Katarina Merganicova,
- Tobias Mette,
- Werner Rammer,
- Kristin H. Braziunas,
- Paola Mairota,
- Josef Bruna,
- Heike Lischke,
- Daniel Scherrer,
- Simon Scheiter,
- Thomas Hickler,
- Tim Anders,
- Jessica Hetzer,
- Harald Bugmann
date: "`r Sys.Date()`"
output:
  bookdown::pdf_document2:
    keep_tex: yes
    number_sections: yes
    toc: no
  word_document:
    reference_docx: template_paper.docx
  bookdown::word_document2: default
mainfont: Arial
fontsize: 12pt
linestretch: 1.5
bibliography: Regeneration.bib
link-citations: yes
linkcolor: blue
csl: draft.csl
editor_options:
  markdown:
    wrap: 72
header-includes:
- \usepackage[left]{lineno}
- \linenumbers
---

@@

Tree regeneration is a key process in forest dynamics, particularly in
the context of forest resilience after disturbance and climate change.
Models of forest dynamics have been in use since more than 50 years, but
an assessment of their capacity to properly represent regeneration is
still lacking. Using 15 models and model variants that were build to
capture long term (decadal to centennial) forest dynamics at the stand,
landscape and global levels, we simulate tree regeneration across at 200
sites representing environmental gradients across Europe. The simulation
results are benchmarked against data from forest reserves collected at
these 200 sites. We aim to assess how well the models of forest dynamics
are capturing the overall abundance, species composition and mortality
in tree recruitment. The results show that most of the models tend to
over-estimate recruitment rates although this is not always connected to
large mortality rates in the early stages of population dynamics. In the
models, the effect of light availability on recruitment rates is
captured better than the effect of temperature and soil moisture, but
the patterns are not consistent across models. Species richness matches
quite well compared to the ranges observed in the data. Models
simulating higher species diversity at the stand level do not feature
higher diversity in recruitment. *Abies alba* was well captured across
temperature and soil moisture gradients by most models, whereas *Fagus
sylvatica* was generally underestimated. *Picea abies* recruitment was
generally over-estimated, particularly at colder sites. HERE TWO
SENCENTECES DISCUSSION/CONCLUSION. HERE A SO WHAT SENTENCE

@@

Fifteen models of forest dynamics (including two model featuring
variants) were used to assess tree regeneration. The selected models
simulate forest dynamics at the stand, landscape and global scale (Table
\@ref(tab:tableModels)). The approaches used for model construction and
their origin differ, with most of the models featuring a largely
"process-based" approach, whereas two models based on formulations
derived from the statistical analysis of inventory data (SYBILA and
xComp). Some of the models largely rely on the original approach
underlying forest gap models (ForClim 1, ForClim 11, LandClim, Landis-II
); whereas some include plant ecophysiological processes (FORMIND,
PICUS, 4C, iLand). These different approaches underlying the models also
has strongly influenced the formulation of tree recruitment.

@@

In most models, there were only small but significant differences in the
species diversity of recruitment between 7 and 10 cm across sites (Table
S2-S4). Furthermore, the observed data and three models (4C, SIBYLA,
Landis II) maintained the recruitment diversity between the 7 and 10 cm
thresholds (they were not significant differences between them, see
Table S2), and the same was visible from the observed data. There are
only few models with large recruitment diversity increasing between 7
and 10 cm (ForCEEPS, TreeMig, LPJ-GUESS).

@@

```{r diff, echo=FALSE, fig.cap="Recruitment rates for the five main species in terms of their basal area share in the observations (top row) and the performance of each model across the environmental gradients (other rows) for the 7 cm recruitment threshold. The values shown are the mean of the 200 samples per site and across the sites in each bins (tile), with 10 bins per gradient. The sizes of the circles represent the ratio between recruited basal area of the species and the total recruited basal area for all species. The absence of a circle indicates a zero basal area share in the recruitment, or the absence of recruitment altogether. The color gradient (for the models only) shows the difference between the simulated and observed ratio of recruited basal area of the species and the total recruited basal area for all the species.", out.width='100%', out.height= '90%'}
knitr::include_graphics(here::here("figures", "HeatEnvGrad_10circles_7.png"))
```

## Mortality in regeneration

@@

The three models with largest overestimation proportion of recruitment
at 7cm threshold showed had the largest mortality between 7 and 10 cm
thershold (Figure \@ref(fig:mortality13)A). The rest of the models that
overestimated recruitment had mortalities lower than the expected shelf
thinning ratio (Figure \@ref(fig:mortality13)B). However, most of the
models with a low variation in mortality ratio between 7 and 10 cm
recruitment (ForCEEPS, ForClim 1, ForClim 11 and LandClim) (Figure S8)
feature a small overestimation of recruitment at the 7 cm threshold
(Figure \@ref(fig:mortality11); cf. Figure S8). Some of the models that
start at a lower number of individuals (7 cm threshold) have a lower
decrease between 7 and 10 cm dbh (ratio 7-10 cm) (Figure S10); in these
models the species competition and self-thinning is either not that
large or it happened before the species reached 7 cm. However, this
pattern it is not consistent across models (Figure S9). There is an
effect of the number of individuals at 7cm on the mortality rate. Most
of the models show a positive effect (increasing recruitment at 7cm
shows increases the mortality between 7 and 10cm). Some models show a
negative effect, however the trend it is not very pronounced and not
always significant regarding the slope of the linear trends per model
(Figure S9 and Table S2).

@@

Accurate simulation of tree recruitment rates is pivotal for projecting
the dynamics of forest stand structure and composition. Most of the
models of forest dynamics evaluated here tended to overestimate tree
recruitment rates across the 200 sites. Overestimating has far-reaching
implications regarding biomass (and carbon) turnover, with a potential
overestimation in their capacity for carbon capture [@pugh2019].
Importantly, excess recruitment requires higher mortality rates in later
stages, either -- as observed for
some models -- between the 7 and 10 cm diameter thresholds, or either
before or
shortly after this stage. Correcting for the expected forest densities
at early
stages is thus equivalent to compensating for a first error (excess
recruitment) by committing a second error (excess mortality). Thus, many
models appear to get the right \'answer\', i.e., correct stand structure
and composition, for the wrong reasons. This is problematic particularly
if the models are to be used under novel conditions (e.g., climate
change; [@huber2021]) or in a decision support context (e.g.,
[@thrippleton2021]).

In spite of this criticism, it remains remarkable that most of the
models were not overly off
with respect to simulated regeneration levels -- after all, few if any
of the models\' establishment routines are sufficiently constrained by
data, and the resulting degrees of freedom that are available for
modeling regeneration (cf. [@konig_tree_2022]) could have implied that
model performance would be much worse.

@@

Light availability as a strong filter for tree regeneration has been
documented widely [@messier1999; @collet2006; @berdanier2016]. However,
it is difficult to collect variables defining light availability on
regeneration at a large spatial and temporal scales. In this study, we
used the total stand basal area as a proxy for light availability. We
found pronounced differences in the stand basal area ranges across
models, but also between models and the observation. This made
impossible to evaluate the recruitment for the extremes of the density
ranges in some models. For example, regeneration rates at low stand
densities are relevant to assess how well forests are recovering e.g.,
after gap creation due to disturbance[@seidl_post-disturbance_2022;
@grubb1977]. We expected that with increasing stand basal area, the
recruitment of the different species would become sparser and drop out
entirely under low-light condition [@klopcic2012; @zell2019]. However,
few models showed this trend, thus indicating that the relation between
recruitment and light availability it is not yet captured correctly in
the models.

@@

Correctly capturing tree species composition of recruitment is important
to assess the future functional diversity of a forest, e.g., its
sensitivity to drought or resilience to disturbances events
[@redmond2015; @seidl_post-disturbance_2022]. In the model simulations,
overall species diversity levels in recruitment were within the observed
ranges. Yet, there were variations on how the main tree species were
represented in the environmental gradient. Most models maintain or even
increase in recruitment diversity between the 7 and 10 cm thresholds,
and the same is visible from the observed data, however the differences
are not significant for all the models. Species richness variations
might be explained by the way each model considers recruitment
[@konig_tree_2022], e.g., in the absence of adults, recruitment can be
linked to either seed rain-dispersal, or to seed/seedling bank, but also
on what processes are formulated with stochastic or deterministic
approach. Differences might also be expected between cohort and
individual trees based models. MAKE REFERENCES TO LITERATURE

Our results do not show differences between models with and without
feedback in recruitment. The adult forests species richness was better
captured than the recruitment species richness, and only few models
overpredicted stand level species richness. Most of the models have
feedback on recruitment from the adult trees, therefore EXPAND ON
FEEDBACK ISSUE

## Mortality in regeneration

As mentioned above, it is reasonable to expect that models that
over-recruit trees will also have a couting-acting high mortality
between the 7 and 10cm thresholds, but this was nit cinsistently evident
from the simulations. This implies that these models must have the
excess of mortality in earlier or later stages. Higher mortality in the
adult component might enhance the share of less shade tolerant species
for those models. There might be a difference between models including
stochastic processes in recruitment and those which do not. There was
also a group of models below the Reineke line indicating a lower
mortality than expected by the self-thinning rule, which may point to
facilitation, or simply a higher mortality rate before the trees reached
7 cm diameter, another possibility would have been that these models
underestimate recruitment which was not the case. This might also be
partly an artifact of the method, as we are expecting space for time
substitution, but in reality we just have one measurement. NEED MORE
REFERENCES AND DISCUSSION POINTS

## Research recomendations

There is also the question of how much level of detail one has to give
to define regeneration processes [@konig_tree_2022; @bugmann2022].
Models with more detailed and complex recruitment modules, will have
also increase parametrisation which will lead to lower generalization as
this values will have to be collected from specific geographical
regions. This is significant relevant when looking at mixed forests and
the species level parametrization. **How this relates to the results?
Future model development recommendations?**

In this study, we considered a wide range of models of forest dynamics,
from stand to global scale. However, the simulation set up was limited
to 1 ha recruitment values. It is important to include spatial
considerations on tree recruitment [@beckage2003] and the lack of it
might represent an more important disadvantage for landscape level
models. Without spatial context we are limiting recruitment to the seed
influx from the stand itself if the model does not have a background
seed input. But even global vegetation models are lacking this dispersal
between cells where there is an abstraction of the space or even the
trees themselves [@hanburybrown_forest_2022]. Landscape level models
have been build to be very accurately at the landscape level. EXPAND ON
THE DIFFERENT SCALE LEVELS AND IMPLICATIONS

Modeling approach seems more important than scale and accuracy of the
design. Model behavior (accuracy) is completely unrelated to model scale
(stand, landscape, global) and approach (empirical, process-based).
Having both empricial and process based models, the expectation was that
the empirically based models (SIBYLA, xComp) would have a better
performance as they are calibrated with inventory data, however this was
not the case. EXPAND ON

The selected models also had limitations on species representation, some
models where not able to simulate certain species due to lack of species
parametrization. We also had one model that simulate PFTs providing a
less flexible comparison the with species values; with the lose of the
species specific information might also represent substantial ecological
representativeness drawbacks [@purves2008].

**METHODS PARAGRAPH**

@@

RECOMENDATION PARAGRAPH

Cautious when using this models to evaluate regeneration post
disturbances we might be overestimation their capacity for carbon
capture

@@

Capture properly recruitment rates, species composition and mortality in
the environmental gradient might impact future forests...

Provide succinct answers to the three questions/objectives stated at the
end of the
Introduction. Perhaps end with a concluding (\"more positive\") final
paragraph.

Models eventually get the right stand structure and composition, most of
them do that
by compensating for one error (overestimation of recruitment) by another
error
(excessive subsequent mortality, if not between 7 and 10 cm, then
shortly
thereafter): \"getting the right answer for the wrong reason\". And this
is
worrysome in general, but particularly (and very acutely) if the models
are to
be used e.g. under changing climatic conditions.

>>>>

---
title: Drivers of tree regeneration remain challenging to be captured by models of
  forest dynamics
author:
- Olalla Díaz-Yáñez,
- Yannek Käber,
- Mats Mahnken,
- Samuel Fischer,
- Xavier Morin,
- Manfred Lexer,
- Christian Hochauer,
- Katarina Merganicova,
- Tobias Mette,
- Werner Rammer,
- Kristin H. Braziunas,
- Paola Mairota,
- Josef Bruna,
- Heike Lischke,
- Daniel Scherrer,
- Simon Scheiter,
- Thomas Hickler,
- Tim Anders,
- Jessica Hetzer,
- Harald Bugmann
date: "`r Sys.Date()`"
output:
  bookdown::pdf_document2:
    keep_tex: yes
    number_sections: yes
    toc: no
  word_document:
    reference_docx: template_paper.docx
mainfont: Arial
fontsize: 12pt
linestretch: 1.5
bibliography: Regeneration.bib
link-citations: yes
linkcolor: blue
csl: draft.csl
editor_options:
  markdown:
    wrap: 72
header-includes:
- \usepackage[left]{lineno}
- \linenumbers
---

@@

Tree regeneration is a key process in forest dynamics, particularly in
the context of forest resilience after disturbance and climate change.
Models of forest dynamics have been in use since more than 50 years, but
an assessment of their capacity to properly represent regeneration is
still lacking. Using 15 models and model variants that were build to
capture long term (decadal to centennial) forest dynamics at the stand,
landscape and global levels, we simulate tree regeneration across at 200
sites representing environmental gradients across Europe. The simulation
results are benchmarked against data from forest reserves collected at
these sites. We aim to assess how well the models of forest dynamics are
capturing the overall abundance, species composition and mortality in
tree recruitment. The results show that most of the models tend to
over-estimate recruitment rates although this is not always connected to
large mortality rates in the early stages of population dynamics. In the
models, the effect of light availability on recruitment rates is
captured better than the effect of temperature and soil moisture, but
the patterns are not consistent across models. Species richness matches
quite well compared to the ranges observed in the data. Models
simulating higher species diversity at the stand level do not feature
higher diversity in recruitment. Looking at the individual species,
*Abies alba* was well captured across temperature and soil moisture
gradients by most models, whereas *Fagus sylvatica* was generally
underestimated. *Picea abies* recruitment was generally over-estimated,
particularly at colder sites. Increasing complexity in the models
regeneration modules did not represent and advantage to better capture
recruitment. Model behavior (accuracy) is completely unrelated to model
scale (stand, landscape, global) and approach (empirical,
process-based). We conclude that even-though the overestimation of
recruitment might represent a challenge for the usability of these
models for projections under uncertain conditions such a climate change
and some model formulations should be reconsidered to correct excessive
recruitment levels, it is still remarkable that all these models can
capture such a complex process like regeneration in relatively correct
ranges while never been parametrized with such a data.

@@

Fifteen models of forest dynamics (including two model featuring
variants) were used to assess tree regeneration. The selected models
simulate forest dynamics at the stand, landscape and global scale (Table
\@ref(tab:tableModels)). The approaches used for model construction and
their origin differ, with most of the models featuring a largely
"process-based" approach, whereas two models based on formulations
derived from the statistical analysis of inventory data (SYBILA and
xComp). Some of the models largely rely on the original approach
underlying forest gap models (ForClim 1, ForClim 11, LandClim, Landis
II); whereas some include plant ecophysiological processes (FORMIND,
PICUS, 4C, iLand). These different approaches underlying the models also
has strongly influenced the formulation of tree recruitment.

@@

In most models, there were only small but significant differences in the
species diversity of recruitment between 7 and 10 cm across sites (Table
S2-S4). Furthermore, the observed data and three models (4C, SIBYLA,
Landis II) maintained the recruitment diversity between the 7 and 10 cm
thresholds (they were not significant differences between them, see
Table S2), the same was visible from the observed data. There are only
few models with large recruitment diversity increasing between 7 and 10
cm (ForCEEPS, TreeMig, LPJ-GUESS).

@@

```{r diff, echo=FALSE, fig.cap="Recruitment rates for the five main species in terms of their basal area share in the observations (top row) and the performance of each model across the environmental gradients (other rows) for the 7 cm recruitment threshold. The values shown are the mean of the 200 samples per site and across the sites in each bins (tile), with 10 bins per gradient. The sizes of the circles represent the ratio between recruited basal area of the species and the total recruited basal area for all species. The absence of a circle indicates a zero basal area share in the recruitment, or the absence of recruitment altogether. The color gradient (for the models only) shows the difference between the simulated and observed ratio of recruited basal area of the species and the total recruited basal area for all the species.", out.width='100%', out.height= '85%'}
knitr::include_graphics(here::here("figures", "HeatEnvGrad_10circles_7.png"))
```

## Mortality in recruitment

@@

The three models with largest overestimation proportion of recruitment
at 7cm threshold showed had the largest mortality between 7 and 10 cm
thershold (Figure \@ref(fig:mortality13)A). The rest of the models that
overestimated recruitment had mortalities lower than the expected shelf
thinning ratio (Figure \@ref(fig:mortality13)B). However, most of the
models with a low variation in mortality ratio between 7 and 10 cm
recruitment (ForCEEPS, ForClim 1, ForClim 11 and LandClim) (Figure S8)
feature a small overestimation of recruitment at the 7 cm threshold
(Figure \@ref(fig:mortality11); cf. Figure S8). Some of the models that
start at a lower number of individuals (7 cm threshold) have a lower
decrease between 7 and 10 cm dbh (ratio 7-10 cm) (Figure S10); in these
models the species competition and self-thinning is either not that
large or it happened before the species reached 7 cm. However, this
pattern it is not consistent across models (Figure S9). There is an
effect of the number of individuals at 7 cm on the mortality rate. Most
of the models show a positive effect (increasing recruitment at 7 cm
shows increases the mortality between 7 and 10 cm). Some models show a
negative effect, however the trend it is not very pronounced and not
always significant regarding the slope of the linear trends per model
(Figure S9 and Table S2).

@@

Accurate simulation of tree recruitment rates is pivotal for projecting
the dynamics of forest stand structure and composition. Most of the
models of forest dynamics evaluated here tended to overestimate tree
recruitment rates across the 200 sites. Overestimating has far-reaching
implications regarding biomass (and carbon) turnover, with a potential
overestimation in their capacity for carbon capture [@pugh2019].
Importantly, excess recruitment requires higher mortality rates in later
stages, either -- as observed for
some models -- between the 7 and 10 cm diameter thresholds, or either
before or
shortly after this stage. Correcting for the expected forest densities
at early
stages is thus equivalent to compensating for a first error (excess
recruitment) by committing a second error (excess mortality). Thus, many
models appear to get the right \'answer\', i.e., correct stand structure
and composition, for the wrong reasons. This is problematic particularly
if the models are to be used under novel conditions (e.g., climate
change; @huber2021) or in a decision support context (e.g.,
@thrippleton2021).

In spite of this criticism, it remains remarkable that most of the
models were not overly off
with respect to simulated regeneration levels -- after all, few if any
of the models\' establishment routines are sufficiently constrained by
data, and the resulting degrees of freedom that are available for
modeling regeneration (cf. @konig_tree_2022) could have implied that
model performance would be much worse.

@@

Light availability as a strong filter for tree regeneration has been
documented widely [@messier1999; @collet2006; @berdanier2016]. However,
it is difficult to collect variables defining light availability on
regeneration at a large spatial and temporal scales. In this study, we
used the total stand basal area as a proxy for light availability. We
found pronounced differences in the stand basal area ranges across
models, but also between models and the observation. This made
impossible to evaluate the recruitment for the extremes of the density
ranges in some models. For example, regeneration rates at low stand
densities are relevant to assess how well forests are recovering e.g.,
after gap creation due to disturbance [@seidl_post-disturbance_2022;
@grubb1977]. We expected that with increasing stand basal area, the
recruitment of the different species would become sparser and drop out
entirely under low-light condition [@klopcic2012; @zell2019]. However,
few models showed this trend, thus indicating that the relation between
recruitment and light availability it is not yet captured correctly in
the models.

@@

Correctly capturing tree species composition of recruitment is important
to assess the future functional diversity of a forest, e.g., its
sensitivity to drought or resilience to disturbances events
[@redmond2015; @seidl_post-disturbance_2022]. In the model simulations,
overall species diversity levels in recruitment were within the observed
ranges. Yet, there were variations on how the main tree species were
represented in the environmental gradient. Most models maintain or even
increase in recruitment diversity between the 7 and 10 cm thresholds,
and the same is visible from the observed data, however the differences
are not significant for all the models. Species richness variations
might be explained by the way each model considers recruitment
[@konig_tree_2022], e.g., in the absence of adults, recruitment can be
linked to either seed rain-dispersal, or to seed/seedling bank, but also
on what processes are formulated with stochastic or deterministic
approach. The recruitment rates were sampled from simulations in the
equilibrium, and in this state it is expected that no climax species
such as *Pinus sylvestris* or *Quercus* would be less represented
[@klopcic2015], and most of the models did capture this difference also
present in the observed data.

The adult forests species richness was better captured than the
recruitment species richness, and only few models overpredicted stand
level species richness. Recruitment composition defined as a
proportional to adult abundance regardless of productivity might be
consider a cautious assumption to safely maintain relative species
abundances [@hanbury-brown2022]. Most of the models have feedback on
recruitment from the adult trees, however we did not see improvements on
the species richness representation compared to those models without
feedback. This might have more to do with the models requirements for a
successfull regeneration than with the habitat niche of the adult trees
[@grubb1977].

## Mortality in recruitment

As mentioned above, it is reasonable to expect that models that
over-recruit trees will also have a cauting-acting high mortality
between the 7 and 10 cm thresholds, but this was not consistently
evident from the simulations. This implies that these models must have
the excess of mortality in earlier or later stages. Higher mortality in
the adult component do have implications on forest dynamics and the
expected products forests provide. On the one hand, mortality in later
stages might enhance the share of less shade tolerant species for those
models [@klopcic2015] and therefore shifting the species composition. On
the other hand high densities for a large period of time might
overestimate the potential role of trees regrowth in carbon sink
dynamics [@pugh2019]. There are several factors that lead to mortality
in recruitment such as competition [@casper1997], but there might be
also a differences in mortality between how models define the process
such as the inclusion of stochasticity. These observations indicate that
better reconcile differences between recruitment levels and larger size
classes densities and better representations of early stage mortality
processes are needed.

## Research limitations and recommendations

There is also the question of how much level of detail one has to give
to define regeneration processes [@konig_tree_2022; @bugmann2022].
Models with more detailed and complex recruitment modules, will have
also increase parametrisation which will lead to lower generalization as
this values will have to be collected from specific geographical
regions. This is significant relevant when looking at mixed forests and
the species level parametrization. Based in our results there was not a
significant improvement on how recruitment was captured with increasing
complexity in the recruitment modules and even some of the more complex
models had more dificulties correctly capturing recruitment.

It is important to include spatial considerations on tree recruitment
[@beckage2003]. In this study, we considered a wide range of models of
forest dynamics, from stand to global scale. However, the simulation set
up was limited to 1 ha recruitment values. The lack of large spatial
considerations might represent a more important disadvantage for
landscape level models as they have been built to be very accurately at
the landscape level. Without spatial context we are limiting recruitment
to the seed influx from the stand itself if the model does not have a
background seed input. But even global vegetation models are lacking
this dispersal between cells where there is an abstraction of the space
or even the trees themselves [@hanburybrown_forest_2022]. However, for
capturing recruitment modeling design seems more important than scale
and accuracy. Model behavior (accuracy) is completely unrelated to model
scale (stand, landscape, global) and approach (empirical,
process-based). Having both empirical and process based models, the
expectation was that the empirically based models (SIBYLA, xComp) would
have a better performance as they are calibrated with inventory data,
however this was not the case.

@@

@@

Capture properly recruitment rates, species composition and mortality in
the environmental gradient might impact future forest. Models of forest
dynamics are important tools to represent forest dynamics. The 15 models
and variants used in this study represent a wide range of approaches in
simulating forest dynamics but also face similar challenges when
representing tree recruitment. Tree recruitment levels were generally
overestimated, and recruitment niche was not always correctly captured
as a function of the biotic (light) and abiotic (temperature and
moisture) [@bugmann2001]. The initial composition of tree populations
simulated by the different models was well represented. Surprisingly
changes in the way trees species are represented in recruitment did not
have a large impact on the recruitment composition , such as the
inclusion of feedback from adults trees. Related to mortality in tree
in-growth, the models with higher over estimation of recruitment also
had a larger tree mortality however this was not enough to reduce the
high recruitment densities.

A concluding (\"more positive\") final paragraph.
